# Markers visualization


First, we will take top 10 ranked genes based in Log FC and visualize their
expression in clusters using a heatmap representation.


```{r, heatmap}
top10 <- pbmc.degs %>% 
             group_by(cluster) %>% 
             top_n(n = 10, wt = avg_log2FC)
DoHeatmap(pbmc.seurat, 
          features = top10$gene) + NoLegend()
```

IL-7 is a marker for naive CD4+ T cells, while GZMB is a marker for CD8 T cells.
Then, we can tentatively consider cluster 0 and 2 as CD4 and CD8 T cells,
respectively. 

We can visualize additional known canonical markers in order to assign cell
categories. 

```{r, umap_markers_vis, fig.height=12}
canonical_markers <- c('IL7R',     ## CD4+ cell
                       'CCR7',     ## Naive CD4+ T cell
                       'CD8A',     ## CD8+
                       'NKG7',     ## NK
                       'MS4A1',    ## B cell marker
                       'FCGR3A',   ## Mono
                       'FCER1A',   ## DC
                       'CD14'     ## Mono
                       )

FeaturePlot(pbmc.seurat, 
            features = canonical_markers,
            ncol = 2)
```

Now, we will annotate the cells with their identified identities in the seurat 
object. We will map the cluster names as follows:


```{r, cell_type_definition}
mapping <- data.frame(seurat_cluster=c('0', '1', '2', '3'),
                      cell_type=c('Lymphocyte', 
                                  'B cell', 
                                  'NK',
                                  'Monocyte'))
mapping
```

To assign the new labels we can use the map function from the plyr R package
as follows:


```{r, mapping}
pbmc.seurat$'cell_type' <- plyr::mapvalues(
  x = pbmc.seurat$seurat_clusters,
  from = mapping$seurat_cluster,
  to = mapping$cell_type
)
```


Now, we can plot the clusters with the assigned cell types.


```{r, cell_types}
DimPlot(pbmc.seurat, 
        group.by = 'cell_type',   ## set the column to use as category
        label = TRUE)  +          ## label clusters
        NoLegend()                ## remove legends
```


